---
title: "ReproAnalysis_CP1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

load data into data frame `data`. 
```{r}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Calculate, Make a histogram, Calculate and report the mean and median of the total number of steps taken per day
```{r} 
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), xlab="Number of Steps")
mean_step <- mean(steps_by_day$steps)
median_step <- median(steps_by_day$steps)
```

The `mean` is `r mean_step` and the `median` is `r median_step`.

## What is the average daily activity pattern?
```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

max_step_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
```

The 5-minute interval, on average across all the days in the data set, containing the maximum number of steps is `r max_step_interval`.

## Impute missing values. 
1. total number of missing values in the dataset (i.e. the total number of rows with NAs) 
```{r}
sum(!complete.cases(data))
```
2. filling in all of the missing values in the dataset with the average step per interval
```{r}
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```
3. Recalculate total steps by day and create Histogram. 
```{r}
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

#Create Histogram to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```
4. Calculate new mean and median for imputed data. 
```{r}
mean_step_i <- mean(steps_by_day_i$steps)
median_step_i <- median(steps_by_day_i$steps)
```

Calculate difference between imputed and non-imputed data.
```{r}
mean_diff <- mean_step_i - mean_step
med_diff <- median_step_i - median_step_i
```

Calculate total difference.
```{r}
total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
Created a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends.  
``` {r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```